---
title: "1. Get started "
author: "Paul Taconet"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
format:
  html:
    code-overflow: scroll
---

```{r, include = FALSE}
NOT_CRAN <- identical(tolower(Sys.getenv("NOT_CRAN")), "true") # vignette will not be executed when tested on the cran
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  purl = NOT_CRAN
  )
library(emo)
```

```{r setup, message=F}
library(modisfast)
library(sf)
library(terra)
```

In this example we want to download, import and plot over the 3500 km^2^ wide region of interest (mapped below) :

- a 1 year-long time series of monthly NDVI from [MODIS Terra Vegetation indices](https://dx.doi.org/10.5067/MODIS/MOD13A3.061) (spatial resolution : 1 km ; temporal resolution : 1 month),
- the same 1 year-long times series of monthly precipitations from [Global Precipitation Measurement (GPM)](https://doi.org/10.5067/GPM/IMERG/3B-MONTH/07) (spatial resolution : 1° ; temporal resolution : 1 month)

```{r, echo=F, fig.height=5, fig.width=3, eval=T }
library(mapview)
roi <- st_as_sf(data.frame(geom="POLYGON ((-5.82 9.54, -5.42 9.55, -5.41 8.84, -5.81 8.84, -5.82 9.54))"),wkt="geom",crs = 4326)
mapview::mapview(roi,legend=F)
```

### Check which collections are available 

Before starting, let's identify our collections of interest among the collections that are available for download, using the function `mf_list_collections()` :

```{r mf_list_collections, eval=T}
collections <- mf_list_collections()
print(str(collections))
```

So our collections of interest are : 

- **MOD13A3.061** : MODIS/Terra Vegetation Indices Monthly L3 Global 1 km SIN Grid
- **GPM_3IMERGM.07** : GPM IMERG Final Precipitation L3 1 month 0.1 degree x 0.1 degree V07

### Setup the region and time range of interest

First we prepare the script : define the ROI, the time frame, and the folder were the data will be downloaded :

```{r example_prepare, eval=T}
# Set ROI and time range of interest
roi <- st_as_sf(data.frame(id="Korhogo", geom="POLYGON ((-5.82 9.54, -5.42 9.55, -5.41 8.84, -5.81 8.84, -5.82 9.54))"), wkt="geom", crs = 4326)
time_range <- as.Date(c("2017-01-01","2017-12-31"))
target_fold <- "env_data"
```

### Login to EOSDIS Earthdata

And we login to EOSDIS Earthdata with our credentials. To create an account go to : https://urs.earthdata.nasa.gov/.

```{r log1, eval=T, echo=F,message=F }
username <- Sys.getenv("earthdata_un")
password <- Sys.getenv("earthdata_pw")
log <- mf_login(credentials = c(username,password),verbose = F)
```

```{r odr_login, eval=F}
# Login to Earthdata servers with username and password. 
username <- "earthdata_un"
password <- "earthdata_pw"
log <- mf_login(credentials = c(username,password))
```

```{r log3, eval=T, echo=F,message=F }
username <- Sys.getenv("earthdata_un")
password <- Sys.getenv("earthdata_pw")
log <- mf_login(credentials = c(username,password),verbose = F)
```

```{r eval=T, echo=F,message=F }
unlink(target_fold, recursive = T)
```


### Get the URLs of the data to download {#get-url}

With the function `mf_get_url()`, we get the https URLs for our collections of interest (*MOD11A1.061* and *GPM_3IMERGDF.07*) given our ROI and time frame. 

`r emo::ji("right arrow")` For each collection, we can select a subset only of variables (sometimes called "dimensions", or "bands") that are of interest for us, through the parameter `variables`. By default, this parameter is set to `NULL`, which means that all the available variables for the specified collection are included. To get the variables available for a given collection along with information for each of them (description, etc.), use the function `mf_list_variables()`.

<!-- `r emo::ji("right arrow")` **Note on the use of the parameter `variables` of the function `mf_get_url()` :** -->

<!-- The collections generally contain several variables (sometimes called "dimensions", or "bands"). As an example, MODIS LST products contain one band for the day temperature ("LST_Day_1km"), one band for the night temperature ("LST_Night_1km"), etc. To get all the variables available for a given collection along with information for each of them (description, etc.), use the function `mf_list_variables()`. e.g. : -->

<!-- ```{r mf_list_variables, eval=T} -->
<!-- tail(mf_list_variables("MOD11A1.061")) -->
<!-- ``` -->

<!-- In `mf_get_url()`, the parameter `variables` enables to restrict the data to download to only specific variables. By default it is set to `NULL`, which means that all the available variables for the specified collection are downloaded. Specifying `variables` will make the data to download lighter.  -->

<!-- **End Note** -->


```{r mf_get_url, eval=T}
## Get the URLs of MODIS Vegetation indexes
urls_mod11a1 <- mf_get_url(
  collection = "MOD13A3.061",
  variables = c("_1_km_monthly_NDVI","_1_km_monthly_EVI","_1_km_monthly_VI_Quality"),  # get the available variables with : mf_list_variables("MOD13A3.061"), or set to NULL (default) to download all the variables
  roi = roi,
  time_range = time_range )

## Get the URLs of GPM daily
urls_gpm <- mf_get_url(
  collection = "GPM_3IMERGM.07",
  variables = c("precipitation"),  # get the available variables with : mf_list_variables("GPM_3IMERGM.07")
  roi = roi,
  time_range = time_range )

nrow(urls_mod11a1)
urls_mod11a1

nrow(urls_gpm)
head(urls_gpm,3)

```

### Download the data {#download}

Now we download the data with the function `mf_download_data()`.

The destination file for each dataset is specified in the column *destfile* of the dataframes *urls_mod11a1* and *urls_gpm*. A destination file is specified by default but it can of course be modified.

Setting the argument `parallel` to `TRUE` will parallelize - therefore fasten - the download in case numerous datasets need to be downloaded.

```{r mf_download_data, eval=T }

# Hint : to fasten the download in case of multiple collections, bind them in a single data.frame (as shown below) before executing the function 'mf_download_data()' :
df_to_dl <- rbind(urls_mod11a1,urls_gpm)

# Then download the data (by default, in temp directory, but the target folder can be set in the function
res_dl <- mf_download_data(df_to_dl, path = target_fold, parallel = TRUE)

print(str(res_dl))
```

### Import the data in R {#import}

Various packages and related classes can be used to read the data downloaded through OPeNDAP. If [`terra`](https://cran.r-project.org/package=terra) (successor of [`raster`](https://cran.r-project.org/package=raster)) is surely the most famous class for raster objects, many packages facilitate the use of spatiotemporal data cubes in formats such as those proposed through `modisfast` (e.g. NetCDF). For instance, MODIS or VIIRS products can be imported as a `stars` object from the excellent [`stars`](https://cran.r-project.org/package=stars) package for data cubes manipulation. All the data can also be imported as `ncdf4` objects using e.g. the [`ncdf4`](https://cran.r-project.org/package=ncdf4) package.

`r emo::ji("warning")` `r emo::ji("warning")` `r emo::ji("warning")` `r emo::ji("warning")` 

#### Important warning regarding the use of the data after download through `modisfast` {#warning-import}

Care must be taken when importing data that was downloaded through the OPeNDAP data providers servers. Depending on the collection, some "issues" were raised. These issues are independent from `modisfast` : they result most of time of a kind of lack of full implementation of the OPeNDAP framework by the data providers. These issues are :

- for MODIS and VIIRS collections : CRS has to be provided
- for GPM collections : CRS has to be provided + data have to be flipped

**End warning**

These issues can easily be dealt at the import phase in R using the function `mf_import_data()`. This function enables to open the data in R as a `SpatRaster` object (from the package [`terra`](https://cran.r-project.org/package=terra)) :

```{r import_example, message=F, eval=T}
# import MOD11A1.061
modis_ts <- mf_import_data(path = file.path(target_fold,"Korhogo","MOD13A3.061"), 
                           collection_source = "MODIS")
# import GPM
gpm_ts <- mf_import_data(path = file.path(target_fold,"Korhogo","GPM_3IMERGM.07"), 
                         collection_source = "GPM")

modis_ts
gpm_ts
```

### Plot the data

Let's finally plot the data !

(Note that only the first 12 dates are plotted here)

```{r plot , fig.fullwidth = TRUE, fig.height = 10, fig.width = 9,fig.align = 'center', eval=T}
# Land surface temperature
terra::plot(modis_ts["_1_km_monthly_NDVI"])
# Precipitation
terra::plot(gpm_ts["precipitation"])
```


