---
title: "Get started !"
author: "Paul Taconet"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.fullwidth = TRUE, 
  fig.height = 10, 
  fig.width = 9,
  fig.align = 'center',
  message = FALSE,
  warning = FALSE
)
```

```{r setup, message=F}
require(opendapr)
require(sf)
require(stars)
require(raster)
require(ncdf4)
require(magrittr)
require(purrr)
```

## Which data are available through the package ? 

The function `get_collections_available()` provides the collections that are available for download through the `opendapr` package

```{r get_collections_available, eval=F}
(collections_available <- get_collections_available())

# to get additional information for each collection : 
# opendapr:::opendapMetadata_internal

```

The collections currently implemented are listed in the following table : 

```{r tab_collections_available, echo=F}
rmarkdown::paged_table(get_collections_available())
```

If we want to get detailed information on a specific collection, we use the function `get_collection_info()`. The function takes as input a single character string argument, the collection identifier, which is the column "collection" of the output data.frame of the function `get_collections_available()`. 

The function `get_collection_info()` will simply open a new browser tab with the metadata of the collection. 

As an example, for the collection MOD11A1.006 : 

```{r get_collection_info, include = TRUE, eval=F}
get_collection_info("MOD11A1.006")
```

## Typical workflow : Download some data over a given region of interest and time range and import the data

Now that we know which collections are available, time to import the data !

Here we take back the example of the readme file, and we import the data in R afterwards.

We want to download over the 3500 km^2^ wide region of interest (mapped below) :

- a 40 days time series of land surface temperature from [MODIS Terra LST](https://dx.doi.org/10.5067/MODIS/MOD11A1.006) (collection *MOD11A1.006*) (spatial resolution : 1 km ; temporal resolution : 1 day),
- the same 40 days times series of precipitations from [Global Precipitation Measurement (GPM)](https://doi.org/10.5067/GPM/IMERGDF/DAY/06) (collection *GPM_L3/GPM_3IMERGDF.06*) (spatial resolution : 1° ; temporal resolution : 1 day)
- the same 40 days times series of soil moisture from [SMAP Daily](https://doi.org/10.5067/T90W6VRLCBHI) (collection *SMAP/SPL3SMP_E.003*) (spatial resolution : 9 km ; temporal resolution : 3 days)

```{r, echo=F, fig.height=5, fig.width=3 }
require(mapview)
roi <- sf::st_read(system.file("extdata/roi_example.gpkg", package = "opendapr"),quiet=TRUE)
mapview::mapview(roi,legend=F)
```

### Prepare the script

First we prepare the script : set-up ROI, time frame and connect to USGS with our credentials

```{r example_prepare}
# Set ROI and time range of interest
roi <- st_read(system.file("extdata/roi_example.gpkg", package = "opendapr"),quiet=TRUE)
time_range <- as.Date(c("2017-01-01","2017-01-30"))

# Login to USGS servers with username and password. To create an account go to : https://ers.cr.usgs.gov/register/
credentials_usgs <- config::get("usgs")
log <- login_usgs(c(credentials_usgs$usr,credentials_usgs$pwd))

```

### Get the OPeNDAP URLs {#get-url}

With the function `get_url()`, we get the OPeNDAP URLs for our collections of interest (*MOD11A1.006*, *GPM_L3/GPM_3IMERGDF.06* and *SMAP/SPL3SMP_E.003*) given our ROI and time frame.

```{r get_url}
## Get the URLs of MODIS Terra LST daily
urls_mod11a1 <- get_url(
  collection = "MOD11A1.006",
  variables = c("LST_Day_1km","LST_Night_1km","QC_Day","QC_Night"),
  roi = roi,
  time_range = time_range
 )

## Get the URLs of GPM daily
urls_gpm <- get_url(
  collection = "GPM_L3/GPM_3IMERGDF.06",
  variables = c("precipitationCal","precipitationCal_cnt"),
  roi = roi,
  time_range = time_range
 )

## Get the URLs of SMAP 3-days
urls_smap <- get_url(
  collection = "SMAP/SPL3SMP_E.003",
  variables = c("Soil_Moisture_Retrieval_Data_AM_soil_moisture",
  "Soil_Moisture_Retrieval_Data_AM_retrieval_qual_flag",
  "Soil_Moisture_Retrieval_Data_PM_soil_moisture_pm",
  "Soil_Moisture_Retrieval_Data_PM_retrieval_qual_flag_pm"),
  roi = roi,
  time_range = time_range
 )

head(urls_mod11a1,3)

head(urls_gpm,3)

head(urls_smap,3)
```

Note that for each collection, we have specified our variables (i.e. dimensions) of interest (argument `variables`). To get the full list of available variables for each collection along with their metadata, we can use the function `get_variables_info()`. As an example :

```{r get_variables_info, eval=F }
var_info_mod11a1 <- get_variables_info("MOD11A1.006")
```

### Download the data {#download}

Now we download the data with the function `download_data()`.

Destination file for each dataset is specified in the column "destfile" of the dataframes *urls_mod11a1* and *urls_gpm*. 

Setting the argument `parallel` to `TRUE` will fasten the download in case their are multiple data to download, because downloading will be parallelized

```{r download_data, eval = FALSE }
df_to_dl <- rbind(urls_mod11a1,urls_gpm,urls_smap)
res_dl <- download_data(df_to_dl,data_source="usgs",parallel = TRUE)
```

### Import the data in R {#import}

We finally import the data in R. Various classes can be used, depending on the collection. For instance, MODIS or VNP products can be imported as `stars` objects, to further use the excellent [`stars`](https://cran.r-project.org/package=stars) package for data cubes manipulation. All the data can also be imported as netcdf datasets using e.g. the [`ncdf4`](https://cran.r-project.org/package=ncdf4) package, or `RasterLayer` / `RasterStackBrick` of the [`raster`](https://cran.r-project.org/package=raster) package.

In any case, care must be taken when importing data that was downloaded through OPeNDAP. Depending on the collection, some "issues" were raised. These issues are independant from `opendapr` : they result most of time of a kind of lack of full implementation of the OPeNDAP framework by the data providers. These issues are :

- for MODIS and VNP collections : CRS is not included
- for GPM collections : data are flipped
- for SMAP collections : data miss information on the bounding coordinates

These issues can easily be dealt at the import phase in R. The script below includes the manipulations that have to be done at the data import phase.

#### Import MODIS / VNP {#import-modis-vnp}

We set-up the modis CRS when we import MODIS or VNP data

```{r import_modis_vnp }
modis_crs <- "+proj=sinu +lon_0=0 +x_0=0 +y_0=0 +a=6371007.181 +b=6371007.181 +units=m +no_defs"

####################################
# Import the MODIS time series as a rasterBrick object (1 rasterBrick for each dimension) :
mod11a1_rast_day <- urls_mod11a1$destfile %>%
  brick(varname="LST_Day_1km",crs=modis_crs)
####################################

# The same can be done with the other variables (LST_Night_1km, etc.)
```

We can also import the same MODIS time series as a `stars` object. Here the interesting point is that all the dimensions are imported at once.

```{r import_modis_vnp_stars }
(mod11a1_stars <- stars::read_stars(urls_mod11a1$destfile) %>% st_set_crs(modis_crs))
# plot(mod11a1_stars)
```

#### Import GPM {#import-gpm}

We flip the raster when we import GPM data 

```{r import_gpm, message=F}
gpm_crs <- "+init=epsg:4326 +proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0 "

####################################
## Import the GPM time series as a rasterBrick object (1 rasterBrick for each dimension) :
gpm_rast_precipitationcal <- urls_gpm$destfile %>%
  purrr::map(~raster(.,varname="precipitationCal",crs=gpm_crs)) %>%
  brick() %>%
  t() %>%
  flip("y") %>%
  flip("x")
####################################

names(gpm_rast_precipitationcal) <- urls_gpm$time_start

# The same can be done with the other variables ("precipitationCal_cnt", etc.)
```

#### Import SMAP {#import-smap}

We set-up the missing bounding coordinates. For this we use the function `get_optional_parameters()` of the `opendapr` package.

```{r import_smap, message=F}
smap_crs <- "+proj=cea +lon_0=0 +lat_ts=30 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0" # EPSG : 6933

####################################
## Import the SMAP time series as a rasterBrick object (1 rasterBrick for each dimension) :
smap_sp_bound <- opendapr::get_optional_parameters(roi = roi, collection = "SMAP/SPL3SMP_E.003")$roiSpatialBound$`1`
  
smap_rast_sm_am <- urls_smap$destfile %>%
  purrr::map(~ncdf4::nc_open(.)) %>%
  purrr::map(~ncdf4::ncvar_get(., "Soil_Moisture_Retrieval_Data_AM_soil_moisture")) %>%
  purrr::map(~raster(t(.), ymn=smap_sp_bound[1], ymx=smap_sp_bound[2], xmn=smap_sp_bound[3], xmx=smap_sp_bound[4], crs=smap_crs)) %>%
  brick()
####################################

names(smap_rast_sm_am) <- urls_smap$time_start

# The same can be done with the other variables (Soil_Moisture_Retrieval_Data_AM_retrieval_qual_flag, etc.)
```

### Plot 

Let's plot everything !

Only the first 16 dates are plotted here-under

```{r plot}
# Land surface temperature
mod11a1_rast_day <- projectRaster(mod11a1_rast_day,crs = "+init=epsg:4326 +proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0 ")
plot(mod11a1_rast_day)
# Precipitation
plot(gpm_rast_precipitationcal)
# Soil moisture
smap_rast_sm_am <- projectRaster(smap_rast_sm_am,crs = "+init=epsg:4326 +proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0 ")
plot(smap_rast_sm_am)
```
