---
title: "1. Get started "
author: "Paul Taconet"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
NOT_CRAN <- identical(tolower(Sys.getenv("NOT_CRAN")), "true") # vignette will not be executed when tested on the cran
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  purl = NOT_CRAN
  )
```

```{r setup, message=F}
require(opendapr)
require(sf)
require(stars)
require(raster)
require(ncdf4)
require(magrittr)
require(purrr)
```

In this example we want to download, import and plot over the 3500 km^2^ wide region of interest (mapped below) :

- a 30 days-long time series of land surface temperature from [MODIS Terra LST](https://dx.doi.org/10.5067/MODIS/MOD11A1.006) (spatial resolution : 1 km ; temporal resolution : 1 day),
- the same 30 days-long times series of precipitations from [Global Precipitation Measurement (GPM)](https://doi.org/10.5067/GPM/IMERGDF/DAY/06) (spatial resolution : 1° ; temporal resolution : 1 day)
- the same 30 days-long times series of soil moisture from [SMAP Daily](https://doi.org/10.5067/T90W6VRLCBHI) (spatial resolution : 9 km ; temporal resolution : 2/3 days)

```{r, echo=F, fig.height=5, fig.width=3, eval=NOT_CRAN }
require(mapview)
roi <- st_as_sf(data.frame(geom="POLYGON ((-5.82 9.54, -5.42 9.55, -5.41 8.84, -5.81 8.84, -5.82 9.54))"),wkt="geom",crs = 4326)
mapview::mapview(roi,legend=F)
```

## Check which collections are available 

Before starting, let's identify our collections of interest among the collections that are available for download, using the function `odr_list_collections()` :

```{r odr_list_collections, eval=NOT_CRAN}
collections <- odr_list_collections()
print(str(collections))
```

So our collections of interest are : 

- *MOD11A1.006*
- *GPM_3IMERGDF.06*
- *SPL3SMP_E.003*

## Setup the region and time range of interest

First we prepare the script : define the ROI and the time frame

```{r example_prepare, eval=NOT_CRAN}
# Set ROI and time range of interest
roi <- st_as_sf(data.frame(geom="POLYGON ((-5.82 9.54, -5.42 9.55, -5.41 8.84, -5.81 8.84, -5.82 9.54))"), wkt="geom", crs = 4326)
time_range <- as.Date(c("2017-01-01","2017-01-30"))
```

## Login to EOSDIS Earthdata

And we login to EOSDIS Earthdata with our credentials. To create an account go to : https://urs.earthdata.nasa.gov/.

```{r odr_login, eval=NOT_CRAN}
# Login to Earthdata servers with username and password. Here we have stored our credentials in local environment variables
username <- Sys.getenv("earthdata_un")
password <- Sys.getenv("earthdata_pw")
log <- odr_login(credentials = c(username,password), source = "earthdata")
```

## Get the OPeNDAP URLs of the data to download {#get-url}

With the function `odr_get_url()`, we get the https URLs for our collections of interest (*MOD11A1.006*, *GPM_3IMERGDF.06* and *SPL3SMP_E.003*) given our ROI and time frame. 

**Note on the use of the parameter `variables` of the function `odr_get_url()` :**

The collections generally contain several variables (sometimes called "dimensions", or "bands"). As an example, MODIS LST products contain one band for the day temperature ("LST_Day_1km"), one band for the night temperature ("LST_Night_1km"), etc. To get all the variables available for a given collection along with information for each of them (description, etc.), use the function `odr_list_variables()`. e.g. :

```{r odr_list_variables, eval=NOT_CRAN}
tail(odr_list_variables("MOD11A1.006"))
```

In `odr_get_url()`, the parameter `variables` enables to restrict the data to download to only specific variables. By default it is set to `NULL`, which means that all the available variables for the specified collection are downloaded. Specifying `variables` will make the data to download lighter. 

**End Note**

In our example, we specify some variables for each collection. 

```{r odr_get_url, eval=NOT_CRAN}
## Get the URLs of MODIS Terra LST daily
urls_mod11a1 <- odr_get_url(
  collection = "MOD11A1.006",
  variables = c("LST_Day_1km","LST_Night_1km","QC_Day","QC_Night"),  # get the variables available with : odr_list_variables("MOD11A1.006") ; or set to NULL (defaults) to download all the variables
  roi = roi,
  time_range = time_range
 )

## Get the URLs of GPM daily
urls_gpm <- odr_get_url(
  collection = "GPM_3IMERGDF.06",
  variables = c("precipitationCal","precipitationCal_cnt"),  # get the variables available with : odr_list_variables("GPM_L3/GPM_3IMERGDF.06")
  roi = roi,
  time_range = time_range
 )

## Get the URLs of SMAP 3-days
urls_smap <- odr_get_url(
  collection = "SPL3SMP_E.003",
  variables = c("Soil_Moisture_Retrieval_Data_AM_soil_moisture","Soil_Moisture_Retrieval_Data_AM_retrieval_qual_flag","Soil_Moisture_Retrieval_Data_PM_soil_moisture_pm","Soil_Moisture_Retrieval_Data_PM_retrieval_qual_flag_pm"),     # get the variables available with : odr_list_variables("SMAP/SPL3SMP_E.003")
  roi = roi,
  time_range = time_range
 )


nrow(urls_mod11a1)
head(urls_mod11a1,3)

nrow(urls_gpm)
head(urls_gpm,3)

nrow(urls_smap)
head(urls_smap,3)
```

## Download the data {#download}

Now we download the data with the function `odr_download_data()`.

Destination file for each dataset is specified in the column *destfile* of the dataframes *urls_mod11a1*,  *urls_gpm* and *urls_smap*. The destination file is specified by default but it can of course be modified.

Setting the argument `parallel` to `TRUE` will parallelize - therefore fasten - the download in case their are numerous datasets to download

```{r odr_download_data, eval=NOT_CRAN }
df_to_dl <- rbind(urls_mod11a1,urls_gpm,urls_smap)
res_dl <- odr_download_data(df_to_dl,source="earthdata",parallel = TRUE)

print(str(res_dl))
```

## Import the data in R {#import}

Various packages and related classes can be used to read the data downloaded through OPeNDAP. If `raster` is surely the most famous class for raster objects, many packages facilitate the use of spatiotemporal data cubes in formats such as those proposed through opendapr (e.g. NetCDF). For instance, MODIS or VIIRS products can be imported as a `stars` object from the excellent [`stars`](https://cran.r-project.org/package=stars) package for data cubes manipulation. All the data can also be imported as `ncdf4` objects using e.g. the [`ncdf4`](https://cran.r-project.org/package=ncdf4) package, or `RasterLayer` of the [`raster`](https://cran.r-project.org/package=raster) package.

### Important note regarding the import of the data in R {#data_import_note}

In any case, care must be taken when importing data that was downloaded through the OPeNDAP data providers servers. Depending on the collection, some "issues" were raised. These issues are independant from opendapr : they result most of time of a kind of lack of full implementation of the OPeNDAP framework by the data providers. These issues are :

- for MODIS and VIIRS collections : CRS has to be provided
- for GPM collections : CRS has to be provided + data have to be flipped
- for SMAP collections : CRS + bounding coordinates of the data have to be provided

These issues can easily be dealt at the import phase in R. Below we propose some functions that include the processings that have to be done at the data import phase in order to open the data as `raster` objects. (argument `destfiles` is the path to a dataset downloaded with opendapr - output of `odr_get_url()$destfile` - and `variable` is the name of a variable to import).

### Functions to import the data in R

#### Import MODIS / VIIRS {#import-modis-vnp}

As a `RasterLayer` object :

```{r import_modis_vnp_raster, eval=NOT_CRAN }
## Functions to import MODIS and VIIRS products as RasterLayer object
# In case the ROI covers one single MODIS tile :
require(raster)
.import_modis_onetile <- function(destfiles,variable){
  rasts <- destfiles %>%
    raster::brick(.,varname=variable,crs="+proj=sinu +lon_0=0 +x_0=0 +y_0=0 +a=6371007.181 +b=6371007.181 +units=m +no_defs")
  return(rasts)
}

# In case the ROI covers multiple MODIS tiles :
.import_modis_moretiles <- function(destfiles,variable){
  rasts <- destfiles %>%
    purrr::map(~raster::brick(.,varname=variable,crs="+proj=sinu +lon_0=0 +x_0=0 +y_0=0 +a=6371007.181 +b=6371007.181 +units=m +no_defs")) %>%
    do.call(merge,.)
  return(rasts)
}
```

As a `stars` object :

We can also import the same MODIS time series as a `stars` object. Here the interesting point is that all the dimensions are imported at once.

```{r import_modis_vnp_stars, eval=NOT_CRAN }
## Functions to import MODIS and VIIRS products as stars object
.import_modis_stars <- function(destfiles){
  stars <- stars::read_stars(destfiles) %>% 
    st_set_crs("+proj=sinu +lon_0=0 +x_0=0 +y_0=0 +a=6371007.181 +b=6371007.181 +units=m +no_defs")
  return(stars)
}
# plot(mod11a1_stars)
```

#### Import GPM {#import-gpm}

As a `RasterLayer` object :

```{r import_gpm, message=F, eval=NOT_CRAN}
## Function to import GPM products as RasterLayer object
require(purrr)
.import_gpm <- function(destfiles,variable){
  rasts <- destfiles %>%
    purrr::map(~raster(., varname = variable,crs = "+init=epsg:4326 +proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0 ")) %>%
    raster::brick() %>%
    raster::t() %>%
    raster::flip("y") %>%
    raster::flip("x")
  return(rasts)
}
```

#### Import SMAP {#import-smap}

As a `RasterLayer` object :

We set-up the missing bounding coordinates. For this we use the function `odr_get_opt_param()`.

```{r import_smap, message=F, eval=NOT_CRAN}
## Function to import SMAP products as RasterLayer object
require(purrr)
require(ncdf4)
require(raster)

smap_sp_bound <- opendapr::odr_get_opt_param(roi = roi, collection = "SPL3SMP_E.003")$roiSpatialBound$`1`

.import_smap <- function(destfiles,variable,smap_sp_bound){
 rasts <- destfiles %>%
   purrr::map(~ncdf4::nc_open(.)) %>%
   purrr::map(~ncdf4::ncvar_get(., "Soil_Moisture_Retrieval_Data_AM_soil_moisture")) %>%
   purrr::map(~raster(t(.), ymn=smap_sp_bound[1], ymx=smap_sp_bound[2], xmn=smap_sp_bound[3], xmx=smap_sp_bound[4], crs="+proj=cea +lon_0=0 +lat_ts=30 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0")) %>%  # EPSG : 6933
   raster::brick()
  return(rasts)
}
```

#### Import the data from our example

So let's import the data from our example, using the functions provided below : 

```{r import_example, message=F, eval=NOT_CRAN}
# import MOD11A1.006
mod11a1_rast_day <- .import_modis_onetile(urls_mod11a1$destfile, "LST_Day_1km")
# import GPM
gpm_rast_precipitationcal <- .import_gpm(urls_gpm$destfile, "precipitationCal")
# import SPL3SMP_E.003
smap_rast_sm_am <- .import_smap(urls_smap$destfile, "Soil_Moisture_Retrieval_Data_AM_soil_moisture",smap_sp_bound)

mod11a1_rast_day
gpm_rast_precipitationcal
smap_rast_sm_am
```

## Plot the data

Let's finally plot the data !

(Note that only the first 16 dates are plotted here-under)

```{r plot , fig.fullwidth = TRUE, fig.height = 10, fig.width = 9,fig.align = 'center', eval=NOT_CRAN}
# Land surface temperature
mod11a1_rast_day <- projectRaster(mod11a1_rast_day,crs = "+init=epsg:4326 +proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0 ")
plot(mod11a1_rast_day)
# Precipitation
names(gpm_rast_precipitationcal) <- urls_gpm$time_start
plot(gpm_rast_precipitationcal)
# Soil moisture
smap_rast_sm_am <- projectRaster(smap_rast_sm_am,crs = "+init=epsg:4326 +proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0 ")
names(smap_rast_sm_am) <- urls_smap$time_start
plot(smap_rast_sm_am)
```
