---
title: "Simple workflow : Download data for a given region of interest and time frame"
author: "Paul Taconet"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simple data import workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  eval=F
)
```

```{r setup, message=F}
require(opendapr)
require(sf)
require(stars)
require(raster)
require(ncdf4)
require(magrittr)
require(purrr)
require(config)
```

In this example we want to download, import and plot over the 3500 km^2^ wide region of interest (mapped below) :

- a 40 days time series of land surface temperature from [MODIS Terra LST](https://dx.doi.org/10.5067/MODIS/MOD11A1.006) (collection *MOD11A1.006*) (spatial resolution : 1 km ; temporal resolution : 1 day),
- the same 40 days times series of precipitations from [Global Precipitation Measurement (GPM)](https://doi.org/10.5067/GPM/IMERGDF/DAY/06) (collection *GPM_L3/GPM_3IMERGDF.06*) (spatial resolution : 1° ; temporal resolution : 1 day)
- the same 40 days times series of soil moisture from [SMAP Daily](https://doi.org/10.5067/T90W6VRLCBHI) (collection *SMAP/SPL3SMP_E.003*) (spatial resolution : 9 km ; temporal resolution : 3 days)

```{r, echo=F, fig.height=5, fig.width=3 }
require(mapview)
roi <- sf::st_read(system.file("extdata/roi_example.gpkg", package = "opendapr"),quiet=TRUE)
mapview::mapview(roi,legend=F)
```

## Setup the region and time range of interest and login to the USGS EROS ERS

First we prepare the script : set-up ROI, time frame and connect to USGS with our credentials

```{r example_prepare}
# Set ROI and time range of interest
roi <- st_read(system.file("extdata/roi_example.gpkg", package = "opendapr"),quiet=TRUE)
time_range <- as.Date(c("2017-01-01","2017-01-30"))

# Login to USGS servers with username and password. To create an account go to : https://ers.cr.usgs.gov/register/. Here we use the config package to store our credentials in a dedicated file locally
credentials_usgs <- config::get("usgs")
log <- login_usgs(c(credentials_usgs$usr,credentials_usgs$pwd))
```

## Get the URLs to download the data {#get-url}

With the function `get_url()`, we get the OPeNDAP URLs for our collections of interest (*MOD11A1.006*, *GPM_L3/GPM_3IMERGDF.06* and *SMAP/SPL3SMP_E.003*) given our ROI and time frame. 

**Note : parameter `variables` of the function `get_url()`**

The collections generally contain several variables (sometimes called "dimensions", or "bands"). As an example, MODIS LST products contain one band for the day temperature ("LST_Day_1km"), one band for the night temperature ("LST_Night_1km"), etc. To get all the variables available for a given collection along with information for each of them (meaning, etc.), use the function `get_variables_info`, e.g. :

```{r get_variables_info}
tail(get_variables_info("MOD11A1.006"))
```

In `get_url()`, the `variables` parameter enables to restrict the data to download to only specific variables. By default it is set to `NULL`, which means that all the available variables for the specified collection are downloaded. Specifying `variables` will make the data to download lighter. 

In our example, we specify some variables for each collection. 

```{r get_url}
## Get the URLs of MODIS Terra LST daily
urls_mod11a1 <- get_url(
  collection = "MOD11A1.006",
  variables = c("LST_Day_1km","LST_Night_1km","QC_Day","QC_Night"),
  roi = roi,
  time_range = time_range
 )

## Get the URLs of GPM daily
urls_gpm <- get_url(
  collection = "GPM_L3/GPM_3IMERGDF.06",
  variables = c("precipitationCal","precipitationCal_cnt"),
  roi = roi,
  time_range = time_range
 )

## Get the URLs of SMAP 3-days
urls_smap <- get_url(
  collection = "SMAP/SPL3SMP_E.003",
  variables = c("Soil_Moisture_Retrieval_Data_AM_soil_moisture","Soil_Moisture_Retrieval_Data_AM_retrieval_qual_flag","Soil_Moisture_Retrieval_Data_PM_soil_moisture_pm","Soil_Moisture_Retrieval_Data_PM_retrieval_qual_flag_pm"),
  roi = roi,
  time_range = time_range
 )

head(urls_mod11a1,3)

head(urls_gpm,3)

head(urls_smap,3)
```

## Download the data {#download}

Now we download the data with the function `download_data()`.

Destination file for each dataset is specified in the column *destfile* of the dataframes *urls_mod11a1*,  *urls_gpm* and *urls_smap*. A destination file is specified by default but it can of course be modified.

Setting the argument `parallel` to `TRUE` will parallelize - therefore fasten - the download in case their are numerous datasets to download

```{r download_data }
df_to_dl <- rbind(urls_mod11a1,urls_gpm,urls_smap)
res_dl <- download_data(df_to_dl,data_source="usgs",parallel = TRUE)

print(str(res_dl))
```

## Import the data in R {#import}

We finally import the data in R. 

`r emo::ji("warning")` Have a look at the section [Important note regarding the further import of the data in R](https://ptaconet.github.io/opendapr/index.html#important-note-import) of the README for additional information and an important note on the data import phase in R ! `r emo::ji("warning")` 

### Import MODIS / VNP {#import-modis-vnp}

```{r import_modis_vnp_raster }
# Import the MODIS time series as a rasterBrick object (1 rasterBrick for each dimension) :
  mod11a1_rast_day <- urls_mod11a1$destfile %>%
  raster::brick(varname="LST_Day_1km",crs="+proj=sinu +lon_0=0 +x_0=0 +y_0=0 +a=6371007.181 +b=6371007.181 +units=m +no_defs")

# The same can be done with the other variables (LST_Night_1km, etc.)
```

We can also import the same MODIS time series as a `stars` object. Here the interesting point is that all the dimensions are imported at once.

```{r import_modis_vnp_stars }
(mod11a1_stars <- stars::read_stars(urls_mod11a1$destfile) %>% st_set_crs("+proj=sinu +lon_0=0 +x_0=0 +y_0=0 +a=6371007.181 +b=6371007.181 +units=m +no_defs"))
# plot(mod11a1_stars)
```

### Import GPM {#import-gpm}

```{r import_gpm, message=F}
## Import the GPM time series as a rasterBrick object (1 rasterBrick for each dimension) :
gpm_rast_precipitationcal <- urls_gpm$destfile %>%
  purrr::map(~raster(.,varname="precipitationCal",crs="+init=epsg:4326 +proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0 ")) %>%
  raster::brick() %>%
  raster::t() %>%
  raster::flip("y") %>%
  raster::flip("x")

# The same can be done with the other variables ("precipitationCal_cnt", etc.)
```

### Import SMAP {#import-smap}

We set-up the missing bounding coordinates. For this we use the function `get_optional_parameters()` of the `opendapr` package.

```{r import_smap, message=F}
## Import the SMAP time series as a rasterBrick object (1 rasterBrick for each dimension) :
smap_sp_bound <- opendapr::get_optional_parameters(roi = roi, collection = "SMAP/SPL3SMP_E.003")$roiSpatialBound$`1`
  
smap_rast_sm_am <- urls_smap$destfile %>%
  purrr::map(~ncdf4::nc_open(.)) %>%
  purrr::map(~ncdf4::ncvar_get(., "Soil_Moisture_Retrieval_Data_AM_soil_moisture")) %>%
  purrr::map(~raster(t(.), ymn=smap_sp_bound[1], ymx=smap_sp_bound[2], xmn=smap_sp_bound[3], xmx=smap_sp_bound[4], crs="+proj=cea +lon_0=0 +lat_ts=30 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0")) %>%  # EPSG : 6933
  raster::brick()

# The same can be done with the other variables (Soil_Moisture_Retrieval_Data_AM_retrieval_qual_flag, etc.)
```

## Plot 

Let's plot the data now !

(Note that only the first 16 dates are plotted here-under)

```{r plot , fig.fullwidth = TRUE, fig.height = 10, fig.width = 9,fig.align = 'center'}
# Land surface temperature
mod11a1_rast_day <- projectRaster(mod11a1_rast_day,crs = "+init=epsg:4326 +proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0 ")
plot(mod11a1_rast_day)
# Precipitation
names(gpm_rast_precipitationcal) <- urls_gpm$time_start
plot(gpm_rast_precipitationcal)
# Soil moisture
smap_rast_sm_am <- projectRaster(smap_rast_sm_am,crs = "+init=epsg:4326 +proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0 ")
names(smap_rast_sm_am) <- urls_smap$time_start
plot(smap_rast_sm_am)
```
