---
title: "2. Get data on several regions or periods of interest at a glance"
author: "Paul Taconet"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{2. Get data on several regions or periods of interest at a glance}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
format:
  html:
    code-overflow: scroll
---

```{r, include = FALSE}
NOT_CRAN <- !identical(tolower(Sys.getenv("NOT_CRAN")), "true") # vignette will not be executed when tested on the cran
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  purl = NOT_CRAN
)
```

## Introduction

In the `vignette("get_started")`, we have imported data : 

* from several collections (MOD13A3.061 and GPM_3IMERGM.07) ;
* over one single region of interest (ROI) ;
* for one single time frame of interest (2017-01-01 to 2017-01-30).

So far so good, but what if we need ***multiple regions of interest***, and / or ***multiple time frames of interest*** ? Those case are likely to happen, for instance :

* multiple time frames of interest : we have spatiotemporal sampling data - e.g. species occurrence - that were collected over a large time frame and we want to study how local past environmental / climatic conditions influence the occurrence ;
* multiple regions of interest : we want to compare two areas in terms of their environmental or climatic conditions.

We could use `for` loops or related stuff to do the job. However, this would not be very optimized. In this vignette, we explain why and we show how to optimize the data import in the case of **multiple regions** or **multiple time periods** of interest. Let's start ! 

## Get data over multiple regions of interest at a glance

It is very easy to get data over multiple regions of interest, as `mf_get_url()` naturally supports the setting up of various ROIs, as shown in the example below :

```{r setup, message=F}
library(modisfast)
library(sf)
library(terra)
library(purrr)
```


```{r, eval=NOT_CRAN }

# Define multiple regions of interest
roi <- st_as_sf(data.frame(id=c("Korhogo","Diebougou"),
                           geom=c("POLYGON ((-5.82 9.54, -5.42 9.55, -5.41 8.84, -5.81 8.84, -5.82 9.54))",
                                   "POLYGON ((-3.62 11.03, -3.13 11.04, -3.11 10.60, -3.60 10.60, -3.62 11.03))"
                                   )),wkt="geom",crs = 4326)
                                   
time_range <- as.Date(c("2017-01-01","2017-01-30"))

# define target folder
target_fold <- "env_data_2"

# and then execute the classical workflow

log <- mf_login(credentials = c(Sys.getenv("earthdata_un"),Sys.getenv("earthdata_pw")))

urls_mod11a1 <- mf_get_url(
        collection = "MOD11A1.061",
        variables = c("LST_Day_1km","LST_Night_1km","QC_Day","QC_Night"),
        roi = roi,
        time_range = time_range)

res_dl <- mf_download_data(urls_mod11a1, path = target_fold, parallel = TRUE)

modis_ts_korhogo <- mf_import_data(file.path(target_fold,"data","Korhogo","MOD11A1.061"), collection = "MOD11A1.061")
modis_ts_diebougou <- mf_import_data(file.path(target_fold,"data","Diebougou","MOD11A1.061"), collection = "MOD11A1.061")

```

```{r eval=NOT_CRAN, echo=F,message=F }
unlink(target_fold, recursive = T)
```

## Get data over multiple time periods of interest at a glance

Here, the things are just a bit more different, as explained in this example :

We first setup the time ranges of interest (and as usual, the ROI)

```{r example_prepare, eval=NOT_CRAN}
roi <- st_as_sf(data.frame(id = "Korhogo", geom="POLYGON ((-5.82 9.54, -5.42 9.55, -5.41 8.84, -5.81 8.84, -5.82 9.54))"),wkt="geom",crs = 4326)

time_ranges <- list(as.Date(c("2016-01-01","2016-01-31")),
                    as.Date(c("2017-01-01","2017-01-31")),
                    as.Date(c("2018-01-01","2018-01-31")),
                    as.Date(c("2019-01-01","2019-01-31")))

target_fold <- "env_data_2"

log <- mf_login(credentials = c(Sys.getenv("earthdata_un"),Sys.getenv("earthdata_pw")))

```

Of course, we could loop over the `mf_get_url()` with the time ranges of interest, and get the data. However, the `mf_get_url()` function does query the OPeNDAP servers each time it is called. This query internally imports various data, including OPeNDAP time, latitude and longitude vectors, and this process takes some time. In case you loop over the function for the same ROI and multiple time frames of interest, it will import again and again the same data, which is quite useless.

Here is where the function `mf_get_opt_param()` comes into the game. For a given collection and ROI, this function queries the OPeNDAP server and retrieves the information that we were mentionning in the previous paragraph. This function is actually run within the `mf_get_url()` function, but its output can also be provided as input parameter `opt_param` of `mf_get_url()`. If `mf_get_url()` is queried multiple times for the same couple {collection, ROI}, it is hence more efficient to pre-compute only once the argument `opt_param` using `mf_get_opt_param()` and to further provide it to `mf_get_url()` within a `for` loop or e.g. a `purrr::map()` function.

**To summarize : when we have multiple time frames of interest, we first execute the function `mf_get_opt_param()`. Then, we execute the function `mf_get_url()`, passing the result of `mf_get_opt_param()` in the parameter `opt_param`.**

First execute the function `mf_get_opt_param()` : 

```{r mf_get_url_multiple_timeframes, eval=NOT_CRAN}

opt_param_mod11a1 <- mf_get_opt_param("MOD11A1.061",roi)

```

Then execute the function `mf_get_url()` passing the argument `opt_param` :

```{r wf, eval=NOT_CRAN}

urls_mod11a1 <- purrr::map_dfr(time_ranges, ~mf_get_url( 
  collection = "MOD11A1.061",
  variables = c("LST_Day_1km","LST_Night_1km","QC_Day","QC_Night"),
  roi = roi,
  time_range = .,
  opt_param = opt_param_mod11a1)
  )

str(urls_mod11a1)

```

Now, download and import the data in R :

```{r dl, eval=NOT_CRAN}

res_dl <- mf_download_data(urls_mod11a1, path = target_fold, parallel = TRUE)

modis_ts <- mf_import_data(path = dirname(res_dl$destfile[1]), collection = "MOD11A1.061")

modis_ts

```

